---
title: "Parsimonious model of genetics/microbiome risk scores"
author: "MS"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment =  NA)
library(dplyr)
library(data.table)
library(foreign)
library(nnet)

library(car)
library(tidyr)
library(tidyverse)
library(haven)
library(tibble)
library(readr)
library(purrr)
library(tidystats)
library(pscl)
library(DescTools)
library(bigstatsr)
```


```{r , include=FALSE}



load( file='Data.test.GMscores.RData')


source('Functions_for_Regs_Taxa_gm_LASSO_PRS.R')

```

## Levels of covariate adjustment

## Minimally adjusted models


### The plan
I think we want to build the model up from most basic to most complex.  The plot would have different facets for Model 1-3, and different line color for omic measure (i think) - PRS, alpha, taxa.  
Ethan thinks the most basic model starts with age, sex, PCs, but from GM perspective, we are interested in what genetics adds to GM data.  So I think will do it both ways.

Here are the steps:
0. covs
1. covs, omic
2. covs,  omic, PRS

Here are the steps for the starting with PCs (pc):
0. covs
1. covs, PRS
2. covs,  omic, PRS


```{r , echo=FALSE}

m1.otu<- 'score.all'
covs.m1<- c('male','AGE_GOLD','income','education', 'relocation_age', 'other', 'Dominican', 'Cuban', 'PuertoRican', 'generation','smoker.current.v1.v2', 'PC1', 'PC2','PC3','PC4','PC5')
step.overall<- steps.func(ancom.selbal.g.test, m1.otu, covs=covs.m1, ancom.selbal.g.test$asthma.current)


overall.res<- combine.res( m1.otu,step.overall$Step0.m1, step.overall$Step1.m1, step.overall$Step2.m1,covs=covs.m1,  ancom.selbal.g.test$asthma.current)



step.pc.overall<- steps.pc.func(ancom.selbal.g.test, m1.otu, covs=covs.m1, outcome=ancom.selbal.g.test$asthma.current)


overall.pc.res<- combine.res.pc(m1.otu,step.pc.overall$Step0.m1, step.pc.overall$Step1.m1, step.pc.overall$Step2.m1,covs=covs.m1, ancom.selbal.g.test$asthma.current)

 
```


```{r , echo=FALSE}

overall.res
overall.pc.res

overall.res$p.add.gm
overall.pc.res$p.add.gm

overall.res$p.add.prs
overall.pc.res$p.add.prs

```





## Levels of covariate adjustment
Set 1: gender_v1_v2 ,AGE_GOLD, smoking 
Set 2: gender_v1_v2 ,AGE_GOLD,  income,education_v1_v2  , smoking
Set 3: gender_v1_v2 ,AGE_GOLD,  income,education_v1_v2, relocation_age, background, generation , smoking  

## Minimally adjusted models


### The plan
I think we want to build the model up from most basic to most complex.  
Ethan thinks the most basic model starts with age, sex, PCs, but from GM perspective, we are also interested in what genetics adds to GM data.  So I think will do it both ways.

Here are the steps:
0. age, sex, smoking, PCS
1. age, sex,  smoking, omic
2. age, sex, smoking,  omic, PRS, PCs

Here are the steps for the starting with PCs (pc):
0. age, sex, PCs
1. age, sex, PCs, PRS
2. age, sex, PCs, PRS, omic


```{r , echo=FALSE}

  
m1.otu<- 'score.all'
# m1.otu<- unique(c(tax.sp.asthma.cur$ANCOM.OTU, tax.sp.asthma.cur$LASSO))
step.overall.sp<- steps.func(ancom.selbal.sp.test, m1.otu,covs=covs.m1, ancom.selbal.sp.test$asthma.current)


overall.res.sp<- combine.res(m1.otu,
                               step.overall.sp$Step0.m1, step.overall.sp$Step1.m1, step.overall.sp$Step2.m1, covs=covs.m1, ancom.selbal.sp.test$asthma.current)


step.pc.overall.sp<- steps.pc.func(ancom.selbal.sp.test, m1.otu,covs=covs.m1, ancom.selbal.sp.test$asthma.current)

overall.pc.res.sp<- combine.res.pc(m1.otu,step.pc.overall.sp$Step0.m1, step.pc.overall.sp$Step1.m1,step.pc.overall.sp$Step2.m1, covs=covs.m1,  ancom.selbal.sp.test$asthma.current)

 
```


```{r , echo=FALSE}

overall.res.sp
overall.pc.res.sp

overall.res.sp$p.add.gm
overall.pc.res.sp$p.add.gm

overall.res.sp$p.add.prs
overall.pc.res.sp$p.add.prs
```

```{r , echo=FALSE}


ancom.selbal.OA<- ancom.selbal.g.test[ancom.selbal.g.test$asthma.current.obese %in% c('Obese asthma', 'Neither')]
ancom.selbal.As<- ancom.selbal.g.test[ancom.selbal.g.test$asthma.current.obese %in% c('Asthma only', 'Neither')]
ancom.selbal.Ob<- ancom.selbal.g.test[ancom.selbal.g.test$asthma.current.obese %in% c('Obese only', 'Neither')]



```



## Obese asthma 

```{r , echo=FALSE}

m1.otu<- 'score.OA'

step.overall<- steps.func(ancom.selbal.OA, m1.otu,covs=covs.m1, ancom.selbal.OA$asthma.current)

overall.res.OA<- combine.res( m1.otu,step.overall$Step0.m1, step.overall$Step1.m1, step.overall$Step2.m1, covs=covs.m1,ancom.selbal.OA$asthma.current)



step.pc.overall<- steps.pc.func(ancom.selbal.OA, m1.otu, covs=covs.m1,ancom.selbal.OA$asthma.current)

overall.pc.res.OA<- combine.res.pc(m1.otu,step.pc.overall$Step0.m1, step.pc.overall$Step1.m1, step.pc.overall$Step2.m1, covs=covs.m1, ancom.selbal.OA$asthma.current)

```


```{r , echo=FALSE}

overall.res.OA
overall.pc.res.OA
rm(m1.otu)

```


## Obese only

```{r , echo=FALSE}

m1.otu<- 'score.Ob'

step.overall<- steps.func(ancom.selbal.Ob, m1.otu,covs=covs.m1,  ancom.selbal.Ob$obese)

overall.res.Ob<- combine.res( m1.otu,step.overall$Step0.m1, step.overall$Step1.m1, step.overall$Step2.m1, covs=covs.m1, ancom.selbal.Ob$obese)



step.pc.overall<- steps.pc.func(ancom.selbal.Ob, m1.otu,covs=covs.m1, ancom.selbal.Ob$obese)

overall.pc.res.Ob<- combine.res.pc(m1.otu,step.pc.overall$Step0.m1, step.pc.overall$Step1.m1, step.pc.overall$Step2.m1, covs=covs.m1, ancom.selbal.Ob$obese)

```


```{r , echo=FALSE}

overall.res.Ob
overall.pc.res.Ob
rm(  m1.otu)
overall.res.Ob$p.add.gm
overall.pc.res.Ob$p.add.gm

```


## Asthma only

```{r , echo=FALSE}

m1.otu<- 'score.As'

step.overall<- steps.func(ancom.selbal.As, m1.otu,covs=covs.m1, ancom.selbal.As$asthma.current)


overall.res.As<- combine.res( m1.otu,step.overall$Step0.m1, step.overall$Step1.m1, step.overall$Step2.m1, covs=covs.m1,ancom.selbal.As$asthma.current)



step.pc.overall<- steps.pc.func(ancom.selbal.As, m1.otu,covs=covs.m1, ancom.selbal.As$asthma.current)

overall.pc.res.As<- combine.res.pc(m1.otu,step.pc.overall$Step0.m1, step.pc.overall$Step1.m1, step.pc.overall$Step2.m1, covs=covs.m1, ancom.selbal.As$asthma.current)

```


```{r , echo=FALSE}

overall.res.As
overall.pc.res.As
# rm(m1.otu)

overall.res.As$p.add.gm
overall.pc.res.As$p.add.gm

overall.res.As$p.add.prs
overall.pc.res.As$p.add.prs

```
# Species


```{r , echo=FALSE}
ancom.selbal.sp.OA<- ancom.selbal.sp.test[ancom.selbal.sp.test$asthma.current.obese %in% c('Obese asthma', 'Neither')]
ancom.selbal.sp.As<- ancom.selbal.sp.test[ancom.selbal.sp.test$asthma.current.obese %in% c('Asthma only', 'Neither')]
ancom.selbal.sp.Ob<- ancom.selbal.sp.test[ancom.selbal.sp.test$asthma.current.obese %in% c('Obese only', 'Neither')]
```

## Obese asthma

```{r , echo=FALSE}


m1.otu<- 'score.OA'

step.overall<- steps.func(ancom.selbal.sp.OA, m1.otu,covs=covs.m1, ancom.selbal.sp.OA$asthma.current)


overall.res.sp.OA<- combine.res( m1.otu,step.overall$Step0.m1, step.overall$Step1.m1, step.overall$Step2.m1, covs=covs.m1, ancom.selbal.sp.OA$asthma.current)



step.pc.overall<- steps.pc.func(ancom.selbal.sp.OA, m1.otu, covs=covs.m1, ancom.selbal.sp.OA$asthma.current)

overall.pc.res.sp.OA<- combine.res.pc(m1.otu,step.pc.overall$Step0.m1, step.pc.overall$Step1.m1, step.pc.overall$Step2.m1, covs=covs.m1, ancom.selbal.sp.OA$asthma.current)
# summary(glm(asthma.current~OTU114+ AGE_GOLD+gender_v1_v2, data=ancom.selbal.sp.OA,family='binomial'))

```


```{r , echo=FALSE}
rm(step.pc.overall, step.overall, m1.otu)
overall.res.sp.OA
overall.pc.res.sp.OA


overall.res.sp.OA$p.add.gm
overall.pc.res.sp.OA$p.add.gm


overall.res.sp.OA$p.add.prs
overall.pc.res.sp.OA$p.add.prs
```


## Obese only

```{r , echo=FALSE}

m1.otu<- 'score.Ob'

step.overall<- steps.func(ancom.selbal.sp.Ob, m1.otu, covs=covs.m1,ancom.selbal.sp.Ob$obese)


overall.res.sp.Ob<- combine.res( m1.otu,step.overall$Step0.m1, step.overall$Step1.m1, step.overall$Step2.m1,covs=covs.m1,  ancom.selbal.sp.Ob$obese)



step.pc.overall<- steps.pc.func(ancom.selbal.sp.Ob, m1.otu,  covs=covs.m1,ancom.selbal.sp.Ob$obese)

overall.pc.res.sp.Ob<- combine.res.pc(m1.otu,step.pc.overall$Step0.m1, step.pc.overall$Step1.m1, step.pc.overall$Step2.m1,covs=covs.m1 , ancom.selbal.sp.Ob$obese)
# summary(glm(asthma.current~OTU114+ AGE_GOLD+gender_v1_v2, data=ancom.selbal.sp.Ob,family='binomial'))

```


```{r , echo=FALSE}
rm(step.pc.overall, step.overall, m1.otu)
overall.res.sp.Ob
overall.pc.res.sp.Ob


overall.res.sp.Ob$p.add.gm
overall.pc.res.sp.Ob$p.add.gm

overall.res.sp.Ob$p.add.prs
overall.pc.res.sp.Ob$p.add.prs
```


## Asthma only

```{r , echo=FALSE}

m1.otu<- 'score.As'

step.overall<- steps.func(ancom.selbal.sp.As, m1.otu, covs=covs.m1, ancom.selbal.sp.As$asthma.current)


overall.res.sp.As<- combine.res( m1.otu,step.overall$Step0.m1, step.overall$Step1.m1, step.overall$Step2.m1,covs=covs.m1, ancom.selbal.sp.As$asthma.current)



step.pc.overall<- steps.pc.func(ancom.selbal.sp.As, m1.otu, covs=covs.m1, ancom.selbal.sp.As$asthma.current)

overall.pc.res.sp.As<- combine.res.pc(m1.otu,step.pc.overall$Step0.m1, step.pc.overall$Step1.m1, step.pc.overall$Step2.m1, covs=covs.m1, ancom.selbal.sp.As$asthma.current)
# summary(glm(asthma.current~OTU114+ AGE_GOLD+gender_v1_v2, data=ancom.selbal.sp.As,family='binomial'))

```


```{r , echo=FALSE}
rm(step.pc.overall, step.overall, m1.otu)
overall.res.sp.As
overall.pc.res.sp.As
overall.res.sp.As$p.add.gm
overall.pc.res.sp.As$p.add.gm

overall.res.sp.As$p.add.prs
overall.pc.res.sp.As$p.add.prs
```

```{r , echo=FALSE}


model.compare.gm<- overall.pc.res.sp %>%
  mutate(level='Species', Outcome='Current') %>%
  bind_rows(overall.pc.res %>%
  mutate(level='Genus', Outcome='Current') )%>%
  bind_rows(overall.pc.res.sp.As %>%
  mutate(level='Species', Outcome='Asthma only' ))%>%
 bind_rows(overall.pc.res.sp.Ob %>%
  mutate(level='Species', Outcome='Obesity only' ))%>%
  bind_rows(overall.pc.res.sp.OA %>%
  mutate(level='Species', Outcome='Obese asthma') )%>%
  bind_rows(overall.pc.res.As %>%
  mutate(level='Genus', Outcome='Asthma only' ))%>%
  bind_rows(overall.pc.res.Ob %>%
  mutate(level='Genus', Outcome='Obesity only' ))%>%
  bind_rows(overall.pc.res.OA %>%
  mutate(level='Genus', Outcome='Obese asthma') )%>%
  arrange(level, Outcome)

model.compare.gm$p.add.gm

model.compare.prs<- overall.res.sp %>%
  mutate(level='Species', Outcome='Current') %>%
  bind_rows(overall.res %>%
  mutate(level='Genus', Outcome='Current') )%>%
  bind_rows(overall.res.sp.As %>%
  mutate(level='Species', Outcome='Asthma only' ))%>%
 bind_rows(overall.res.sp.Ob %>%
  mutate(level='Species', Outcome='Obesity only' ))%>%
  bind_rows(overall.res.sp.OA %>%
  mutate(level='Species', Outcome='Obese asthma') )%>%
  bind_rows(overall.res.As %>%
  mutate(level='Genus', Outcome='Asthma only' ))%>%
  bind_rows(overall.res.Ob %>%
  mutate(level='Genus', Outcome='Obesity only' ))%>%
  bind_rows(overall.res.OA %>%
  mutate(level='Genus', Outcome='Obese asthma') )

model.compare.prs

```


```{r , echo=FALSE}


save(model.compare.gm,model.compare.prs,
     file='ModelCompare_R2_GMPRS_score_test_M3.RData')


```







