---
title: "Feature Selection"
author: "MS"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Features Selection Approach

Coda-lasso is computationally efficient and suitable when the focus is the identification of the most associated microbial taxa.
Selbal stands out when the goal is to obtain a parsimonious model with optimal prediction performance, but it is computationally greedy
https://malucalle.github.io/Microbiome-Variable-Selection/
system('git clone https://github.com/UVic-omics/CoDA-Penalized-Regression')
   https://github.com/malucalle/CoDA-Penalized-Regression
   
   
```{r , echo=F}
# system('git clone https://github.com/UVic-omics/CoDA-Penalized-Regression')
# getwd()

```


```{r , echo=FALSE}
# when you have already git clone the repository
# system('git pull https://github.com/UVic-omics/CoDA-Penalized-Regression')
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r , include=FALSE}
# devtools::install_github(repo = "malucalle/selbal")

library(knitr) # rbookdown, kable
library(glmnet) # glmnet
library(selbal) # selbal
library(ggplot2) # draw selbal
library(gridExtra) # grid.arrange
# library(UpSetR) # upset
library(ggforce) # selbal-like plot
library(grid) # grid.draw
# source coda-lasso functions
source(file = '~/CoDA-Penalized-Regression/R/functions.R')
source(file = '~/CoDA-Penalized-Regression/R/functions_coda_penalized_regression.R')
library(tibble)
library(dplyr)
library(data.table)
library(phyloseq)
library(microbiome)  
```

## SelBal

https://evayiwenwang.github.io/Microbiome_variable_selection_tutorial/selbal.html


```{r Data Prep, include=FALSE}

load('sp.sub.RData')
load('BDC-SOL-GOLD-train.RData')

```





## SElBal

## Adjustment variables



```{r , include=T}

covs.3<- dplyr::select(gold,  gender_v1_v2 ,AGE_GOLD,  income,education, relocation_age, background_5, generation,smoker.current.v1.v2 )


 
```


### Adjusted Species level analysis 


```{r , echo=FALSE}


selbal_sp_cv3<- selbal.cv(x = sp.sub, y = gold$asthma.current,  covar=covs.3,
                       logit.acc = 'AUC')

```


```{r , echo=FALSE}
save( selbal_sp_cv3,  file='selbal_sp_current_m3.RData')

```







## SElBal
```{r obese/asthma categories}
OA<- gold[gold$asthma.current.obese %in% c('Neither','Obese asthma'),]
As<- gold[gold$asthma.current.obese %in% c('Neither','Asthma only'),]
Ob<- gold[gold$asthma.current.obese %in% c('Neither','Obese only'),]

sp.sub.OA<- sp.sub %>% dplyr::filter(rownames(.) %in% OA$goldID)
sp.sub.As<- sp.sub %>% dplyr::filter(rownames(.) %in% As$goldID)
sp.sub.Ob<- sp.sub %>% dplyr::filter(rownames(.) %in% Ob$goldID)

   
```

## Adjustment variables

### Adjusted Species level analysis 


#### Obese Asthma

```{r Data Prep, include=T}


covs.3<- dplyr::select(OA,  gender_v1_v2 ,AGE_GOLD,  income,education, relocation_age, background_5, generation,smoker.current.v1.v2 )

(start<- Sys.time())


# 
(total<- Sys.time()-start)

selbal_OA_cv3<- selbal.cv(x = sp.sub.OA, y = OA$asthma.current,  covar=covs.3,
                        logit.acc = 'AUC')
#  # The optimal number of variables is: 2 
# 

(total<- Sys.time()-start)

```




#### Asthma only

```{r Data Prep, include=T}

covs.3<- dplyr::select(As,  gender_v1_v2 ,AGE_GOLD,  income,education, relocation_age, background_5, generation,smoker.current.v1.v2 )




```



```{r M3, include=T}


selbal_As_cv3<- selbal.cv(x = sp.sub.As, y = As$asthma.current,  covar=covs.3,
                        logit.acc = 'AUC')
#  # The optimal number of variables is: 2 
# 

(total<- Sys.time()-start)

```



#### Obese only

```{r , include=T}

covs.3<- dplyr::select(Ob,  gender_v1_v2 ,AGE_GOLD,  income,education, relocation_age, background_5, generation,smoker.current.v1.v2 )


# 

selbal_Ob_cv3<- selbal.cv(x = sp.sub.Ob, y = Ob$obese,  covar=covs.3,
                        logit.acc = 'AUC')

(total<- Sys.time()-start)


```





```{r , echo=FALSE}


save(  selbal_Ob_cv3, 
      selbal_As_cv3, 
      selbal_OA_cv3, 
     file='selbal_sp_multi_current_cv_train_m3.RData')

```